\subsection{Gravitation}

We will simulate the solar system with only the gravitational force affecting the planets.  Newton's gravitational law is stated in equation (\ref{eq:newton}). Where G is the gravitational constant ($6.67 \cdot 10^{-11} \rm{Nm^2/s^2}$), r is the distance between the planets, m is the mass of the object and M is the mass of the other object.

\begin{align}
	\vec{F_G}  =\frac{GmM}{\vec{|r|}^3}\vec{r}
	\label{eq:newton}
\end{align}

Thankfully for us the force is an attractive force. It is worth noting that if the sun is at origo distance is simply the norm position vector of the planet. From now on we will denoted the different masses with $M_{planet name}$ except for the sun that has the special symbol $M_{\odot}$. 
\\
\\
If you have n object attracting each other the total gravitational force, $F_k$, for each object is: 


\begin{align}
	F_k = 
	\sum_{i = 1}^{N}
	\vec{F_i}  
	=
	\frac{Gm_km_i}
	{\vec{|r_k - r_i|}^3}
	(\vec{r_k} - \vec{r_i})
	(1 - \delta_{k,i})
	\label{eq:newton_all}
\end{align}

Where $\delta_{k,i}$ is the function:

\begin{align*}
	\delta_{k,i} = \left\{\begin{matrix}
					1 & \text{if} \quad k =  i\\
					0 & \text{if} \quad k \neq i 
					\end{matrix}\right.
\end{align*}



We can use Newton's second law to determine the accelration of the planet. Newton's second law state $F = m\vec{a}$. Which translate to:

\begin{align}
	a_k
	=
	\sum_{i = 1}^{N}
	\frac{Gm_i}
	{|\vec{r_k} - \vec{r_i}|^3}
	(\vec{r_k} - \vec{r_i})
	(1 - \delta_{k,i})
	\label{eq:acceleration_all}
\end{align}



\subsection{Units}\label{sec:units}
As a famous person once said, \"If you use seconds and meters you will not finish within the deadline\". I think our professor, Morten Hjorth-Jensen, is on to something. Seconds and meters are impractical in this project. Therefore we use more suitable units. The distance between the earth and sun, 149 597 870 691 meter, is defined as one astronomical unit (au). For time one year seems reasonable. 
These are the units that will be used in this project. 
\\
\\
We can now express the gravitational constant with these units. To do this we use the formula for the acceleration for a circular orbit is $\frac{v^2}{r}$. Combine this with the gravitational law, we get: 

\begin{align*}
	\frac{v^2}{r} =
	\frac{GM_{\odot}}{\vec{|r|}^3}\vec{r} 
	\implies 
	G
	=
	\frac{v^2 r}{M_{\odot}} = \frac{\rm{au} (2\pi \cdot \rm{au/year})^2}{M_{\odot}}
	=
	\frac{4\pi^2}
	{M_{\odot}} \rm{au^3/year^2}
\end{align*}

Applying this to equation (\ref{eq:acceleration_all}), we get: 

\begin{align}
	a_k
	=
	\sum_{i = 1}^{N}
	4 \pi \frac{m_i}{M_{\odot}}
	\frac{(\vec{r_k} - \vec{r_i})}
	{|\vec{r_k} - \vec{r_i}|^3}
	(1 - \delta_{k,i}) \rm{au^3/year^2}
	\label{eq:acceleration_all_au}
\end{align}


\subsubsection{Escape velocity}\label{sec:escape-velocity}

Later in the project we are asked to find the escape velocity by trial and error. This question is basicly, when does the earth have enough kinetic energy to escape the potential. The potential is a integral from 0 to $\infty$ $F_G$ over $dr$. For a given planet with only the sun in the solar system the integral is from r to $\infty$. This is shown in equation (\ref{eq:escape-velocity}).  


\begin{align}
E_k =\int_r^\infty \frac{GmM}{\vec{|r|}^2} d\vec{r}\\
\label{eq:escape-velocity}
\end{align}

Let's solve it: 

\begin{align*}
	&E_k =\int_r^\infty -\frac{GmM_\odot}{\vec{|r|}^2} d\vec{r}
	\\
	&\frac{1}{2}mv^2 = GmM_\odot \int_r^\infty -\frac{1}{\vec{|r|}^2} d\vec{r}
	\\
	&\text{G = 1, see section \ref{sec:units} for explanation.}
	\\
	&\frac{1}{2}v^2 = M_\odot \int_r^\infty -\frac{1}{\vec{|r|}^2} d\vec{r}
	\\
	&\frac{1}{2}v^2 = M_\odot \left[ \frac{1}{\vec{|r|}} \right]_r^\infty
	\\
	&v^2 = 2M_\odot \frac{1}{\vec{|r|}}
	\\
	&\text{r = 1, see section \ref{sec:units} for explanation.}
	\\
	&v = \sqrt{8\pi^2}
	\\
	&v \approx 8.88
\end{align*}

\todo{skal jeg ikke dele uttrykket på 3? men dette gir rett svar}







\subsection{prehelion}


\todo{since this is not implemented in the code yet nothing is writen}
















\subsection{Numerical methods}

Equation (\ref{eq:acceleration_all_au}) initial conditions for velocity and position determines the orbit of the planets. 
These equations are for x,y,z direction and when writen out you get a coupled set of differential equations. This set is near impossible to solve analytically. It might be impossible. We will use numerical methods to solve this set. More specifically we will use the Forward Euler method and the Velocity-Verlet method. A given $t_i$ is equal to $t_0 + ih$, where $h = (t_{0} + t_{n})/n $.

\subsubsection{Forward Euler}

Both method use a Taylor polynomial approximation to solve the set of differential equation. The Forward Euler use the first order Taylor polynomial. With $r'(t) = v(t)$ and $v'(t) = a(t)$ the Forward Euler method result in: 
 
\begin{align}
	&\vec{r_i}(t+h) \approx \vec{r_i}(t) + h \vec{v_i}(t)
	\\
	&\vec{v_i}(t+h) \approx \vec{v_i}(t) + h \vec{a_i}(t)
\end{align}

Discretized version. i is the object number and j is the time step:

\begin{align*}
	&\vec{r_{i,j+1}} \approx \vec{r_{i,j}} + h \vec{v}_{i,j}
	\\
	&\vec{v_{i,j+1}} \approx \vec{v_{i,j}} + h \vec{a}_{i,j}
\end{align*}

The error for a first order taylor polynomial goes as $\mathcal{O}(h^2)$ \todo{REF TO LECTURE NOTES}. This is the error for each step. The error is accumulated for each step and will thereby be proportional to h. 

\subsubsection{Velocity-Verlet}

You guessed it, this is the second order Taylor polynomial. . As I said the Velocity-Verlet method is based on a Taylor polynomial approximation. And this is the second order. 

\begin{align}
	&\vec{r_i}(t+h) \approx \vec{r_i}(t) + h \vec{v_i}(t) + \frac{1}{2} h^2 a(t)
	\\
	&\vec{v_i}(t+h) \approx \vec{v_i}(t) + h \vec{a}(t) + \frac{1}{2} h^2 \vec{a}'(t)
	\label{eq:verlet}
\end{align}

Since we have no formula for the derivative of a, we use the approximation: 
\todo{ER DET DELT PÅ 2 ELLER h? } 

\begin{align*}
	\vec{a}'(t) \approx \frac{\vec{a}(t+h) - \vec{a}(t)}{h}
\end{align*}

We simply update the position first. We keep the old acceleration and calculate the acceleration at the new position.  Using this equation (\ref{eq:verlet}) we get: 

\begin{align*}
	&\vec{v_i}(t+h) \approx \vec{v_i}(t) + h \vec{a}(t) + \frac{1}{2} h(\vec{a}(t+h) - \vec{a}(t))
	\\
	&\vec{v_i}(t+h) \approx \vec{v_i}(t) + \frac{1}{2} h(\vec{a}(t+h) + \vec{a}(t))
\end{align*}

Discretized version. i is the object number and j is the time step:
\begin{align*}
	\\
	&\vec{r}_{i,j+1} \approx \vec{r}_{i,j} + h \vec{v}_{i,j} + \frac{1}{2} h^2 a_{i,j}
	\\
	&\vec{v}_{i,j+1} \approx \vec{v}_{i,j} +\frac{1}{2} h(\vec{a}_{i,j+1} + \vec{a}_{i,j})
\end{align*}


The error for a first order taylor polynomial goes as $\mathcal{O}(h^2)$ \todo{REF TO LECTURE NOTES}. This is the error for each step. The error is accumulated for each step and will thereby be proportional to $h^2$. This is one order of magnitude better, then the Forward Euler method. 
















